---
title: "Cleaning and Visualizations"
execute:
 echo: false
 warning: false
embed-resources: true
---

```{r}
#| label: setup

library(tidyverse)
library(janitor)
library(viridisLite)
 
```

```{r}
#| label: data-read-in
dc<-read_csv("data_raw/Esports Study Data - Data Collection.csv", skip=2)
pc<-read_csv("data_raw/Esports Study Data - Participant Contact.csv", skip=3)
pc <- pc |>
  clean_names()
dc <- dc |>
  clean_names()
```

```{r}
#| label: clean-data

junk <- c( #filter for rows
  "Code", "ID Participant Code", "% of total",
  "Mean", "SD", "EM",
  "Casual SSBU", "Elite SSBU", "Comp SSBU",
  "MOBA", "FPS")

drop_cols <- c( #the filter for columns
  "participant_name",
  "email",
  "phone_number",
  "date_for_follow_up_by",
  "m_f",
  "sbp_mm_hg", 
  "dbp_mm_hg",
  "vo2peak_ml_kg_min", 
  "day_1_pro_g", 
  "mean_pro_kg",
  "day_1_cho_g", 
  "mean_cho_kg", 
  "day_1_fat_g",
  "mean_fat_kg", 
  "day_1_fiber_g", 
  "fiber_kg",
  "total_kcal", 
  "total_kcal_kg")

dc_clean <- dc |> #cleans data collection
  filter(
    !is.na(code),
    !code %in% junk,
    !str_detect(code, "^(MOBA|FPS)$")) # only exact MOBA and FPS without numbers after

pc_clean <- pc |> #cleans participant contact
  filter(
    !is.na(id_participant_code),
    !id_participant_code %in% junk,
    !str_detect(id_participant_code, "^(MOBA|FPS)$"))

joined <- inner_join( #joins data as long as the match 
  pc_clean,
  dc_clean,
  by = c("id_participant_code" = "code"))

joined <- joined |>
  mutate(
    # Rename m_f to standardize to F / M
    sex = case_when(
      grepl("^f", m_f, ignore.case = TRUE) ~ "F",
      grepl("^m", m_f, ignore.case = TRUE) ~ "M",
      TRUE ~ NA_character_),
    # Standardize dominant_hand to R / L / Ambi
    dominant_hand = case_when(
      str_detect(dominant_hand, regex("^r", ignore_case = TRUE)) ~ "R",
      str_detect(dominant_hand, regex("^l", ignore_case = TRUE)) ~ "L",
      TRUE ~ NA_character_))|>
  select(
    -matches("^x\\d+$"),
    -any_of(drop_cols))|>
  filter(!is.na(dominant_hand), !is.na(sex))
```

```{r}
#| label: save_and_read_data
write.csv(joined, "data_clean/clean_data.csv", row.names = FALSE)
clean_data <- read_csv("data_clean/clean_data.csv")
```

## Data Cleaning  

```         
Cleaning the data
```

> Row Filtering: Removed non-participant rows (e.g., “Mean”, “SD”, summary labels) to ensure only valid observations were included.
>
> Data Joining: Merged two data sets using participant IDs to combine behavioral and demographic information.
>
> Column Selection: Dropped unnecessary columns (e.g., extra “x#” variables and personal identifiers) and retained only relevant variables up to delta_building_test.

## Visualization 1  

Here is a description of the first visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-1
#| fig-alt: "Scatterplot of Dexterity Building Test scores on the y axis by simple reaction time on the x axis, there are two line of best fits, one for each sex. The two lines are also colored by sex as well as the datapoints being a different shape for each sex"
#| fig-cap: "Scatterplot of Dexterity by Cognition"
#| fig-width: 6 
#| fig-height: 4 
#| fig-align: center
ggplot(clean_data, aes(x = actual_mean_rt_ms, 
                       y = actual_building_test_score, 
                       color = sex, 
                       shape = sex)) +
  geom_point() +
  scale_color_brewer(palette="Pastel1")+
  labs(title = "Dexterity Scores by Cognition Score and Sex ",
       x = "Mean Simple Reaction Time",
       y = "Building Test Score") +
  geom_smooth(method="lm", se=FALSE)+
  theme_minimal()
```

## Visualization 2  

Here is a description of the first visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-2
#| fig-alt: y
#| fig-cap: s
#| fig-width: 6 
#| fig-height: 4 
#| fig-align: center
ggplot(clean_data, aes(x = delta_building_test, 
                       y=sex, 
                       fill = sex, 
                       shape = sex)) +
  geom_violin(alpha=.2)+
  geom_boxplot(alpha=.9)+
  scale_fill_brewer(palette="Pastel1")+
  labs(title = "Dexterity Scores by Sex ",
       x = "Difference in Practice to Actual Building Test Score",
       y = "Sex") +
  theme_minimal()
```
